### 2023-05-04 doc/poster_eccb22_barcelona/poster_eccb22_barcelona.tex
DOC presentation

	%\documentclass[landscape,a0b,final,a4resizeable]{a0poster}
	%\documentclass[landscape,a0b,final]{a0poster}
	%\documentclass[portrait,a0b,final,a4resizeable]{a0poster}
	\documentclass[portrait,a0b,final,a4resizeable]{a0poster}
	%%% Option "a4resizeable" makes it possible ot resize the
	%   poster by the command: psresize -pa4 poster.ps poster-a4.ps
	%   For final printing, please remove option "a4resizeable" !!

@@

	\begin{enumerate}
	\item Genome-wide association studies (GWAS) map genotypes to traits.
	\item Here we use this GWAS database: The IEU OpenGWAS project\\ \url{https://gwas.mrcieu.ac.uk}
	\item The large number of available GWAS have allowed to define pleiotropic regions in the genome (\citep{2019.Posthuma.Watanabe}).
	\item Most GWAS loci are non-coding but the gene regulatory basis of pleiotropy is not well understood.
	\item Expression quantitative trait loci (eQTL) map genotypes to expression phenotypes.
	\item Here we use this eQTL database: The EBI eQTL Catalogue\\ \url{https://www.ebi.ac.uk/eqtl}
	\item The R package coloc performs colocalization tests between genetic traits:\\ \url{https://CRAN.R-project.org/package=coloc}
	\end{enumerate}

@@

	- A pipeline of eQTL/GWAS variant colocalization (Screenshot pythonanywhere)
	- Table with most pleiotropic variants (Table 1)
	- Molecular mechanisms Fig 7 and fig 8a
	- Model of mechanisms Fig 10

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% FIGURE 2 downstream regulation
	\begin{minipage}{1\columnwidth}
	    \begin{center}
		    \begin{minipage}{\columnwidth} 
\includegraphics[width=\textwidth]{img_tmp/david_pleio_3.eps}
\includegraphics[width=\textwidth]{img_tmp/david_pleio_2.eps}
		    \end{minipage}
		    \mycaptionfig{Gene ontology analysis of eGenes of variants associated to three (\textbf{a}) and two (\textbf{b}) GWAS categories compared to eGenes of variants associated to one GWAS category.}
	    \end{center}
	\end{minipage}

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% FIGURE 1 VEP
	\begin{minipage}{1\columnwidth}
	    \begin{center}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/missense_variant.eps}
		    \end{minipage}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/splice_region_variant.eps}
		    \end{minipage}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/3_prime_UTR_variant.eps}
		    \end{minipage}
		    \mycaptionfig{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: Missense variant (Left), splice region variant (Center) and 3’ UTR variant (Right)}
	    \end{center}
	\end{minipage}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% FIGURE 2 downstream regulation
	\begin{minipage}{1\columnwidth}
	    \begin{center}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_etissue_y_egene.eps}
		    \end{minipage}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_egene_y_etissue.eps}
		    \end{minipage}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_egene_etissue_y_gwas.eps}
		    \end{minipage}
		    \mycaptionfig{Count of eGenes per variant-tissue (\textbf{Left}), eTissues per variant-eGene (\textbf{Center}) and GWAS categories per variant-eGene-eTissue (\textbf{Right}).}
	    \end{center}
	\end{minipage}

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% FIGURE 4 Model
	\begin{minipage}{1\columnwidth}
	    \begin{center}
		    \begin{minipage}{1\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/graphical_summary.eps}
		    \end{minipage}
		    \mycaptionfig{Model of regulatory variant pleiotropy. I have investigated
	three possible mechanisms of pleiotropy. Left, Pleiotropic variants have more
	eGenes that result in more functions and more phenotypes. This might arise
	from an enrichment of pleiotropic variants in splicing or 3’ UTR regions.
	Center, I have also found that eGenes of pleiotropic variants are active
	in more etissues which result in more GWAS phenotypes. This might be
	explained from variants being bound by more transcription factors. Left
	Triplets of variant-eGene-eTissues are associated with more GWAS pheno-
	types, which directly affect the number of GWAS phenotypes. I have found
	that this might be explained by an enrichment of missense alleles.}
	    \end{center}
	\end{minipage}

@@

	\begin{itemize}	
	\item We have developed a pipeline of eQTL/GWAS colocalization (Fig. 1).
	\item These colocalization results are in an easy to use web site (Fig. 1).
	\item We have computed a list of pleiotropic eQTLs (Table 1).
	\item Pleiotropic variants are enriched in missense and splice region VEP effect consequences (Fig. 2).
	\item Pleiotropic eQTLs have more eGenes per variant-eTissue, are active in more eTissues per variant-eGene and are associated to more GWAS categories per variant-eGene-eTissue (Fig. 3).
	\item Pleiotropic variants bind transcription factors (Fig. 4).
	\item Model explaining the reason of pleiotropy for eQTLs (Fig. 5).
	\end{itemize}

>>>>

	%\documentclass[landscape,a0b,final,a4resizeable]{a0poster}
	%\documentclass[landscape,a0b,final]{a0poster}
	%\documentclass[portrait,a0b,final,a4resizeable]{a0poster}
	\documentclass[portrait,a0b,final]{a0poster}
	%%% Option "a4resizeable" makes it possible ot resize the
	%   poster by the command: psresize -pa4 poster.ps poster-a4.ps
	%   For final printing, please remove option "a4resizeable" !!

@@

	\begin{enumerate}
	\item Genome-wide association studies (GWAS) map genotypes to traits.
	\item Here we use this GWAS database: The IEU OpenGWAS project\\ \url{https://gwas.mrcieu.ac.uk}
	\item The large number of available GWAS have allowed to define pleiotropic regions in the genome \citep{2019.Posthuma.Watanabe}.
	\item Most GWAS loci are non-coding but the gene regulatory basis of pleiotropy is not well understood.
	\item Expression quantitative trait loci (eQTL) map genotypes to expression phenotypes.
	\item Here we use this eQTL database: The EBI eQTL Catalogue\\ \url{https://www.ebi.ac.uk/eqtl}
	\item Databases of GWAS and eQTL studies allow large colocalization analysis \citep{2021.Li.Mu}.
	\item The R package coloc performs colocalization tests between genetic traits:\\ \url{https://CRAN.R-project.org/package=coloc}
	\end{enumerate}

@@

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% FIGURE 2 downstream regulation
	\begin{minipage}{1\columnwidth}
	    \begin{center}
		    \begin{minipage}{\columnwidth} 
\includegraphics[width=\textwidth]{img_tmp/david_pleio_3.eps}
\includegraphics[width=\textwidth]{img_tmp/david_pleio_2.eps}
		    \end{minipage}
		    \mycaptionfig{Gene ontology analysis of genes associated with eQTLs (eGenes) of variants associated to three (\textbf{Upper}) and two (\textbf{Bottom}) GWAS categories compared to eGenes of variants associated to one GWAS category.}
	    \end{center}
	\end{minipage}

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% FIGURE 1 VEP
	\begin{minipage}{1\columnwidth}
	    \begin{center}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/missense_variant.eps}
		    \end{minipage}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/splice_region_variant.eps}
		    \end{minipage}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/3_prime_UTR_variant.eps}
		    \end{minipage}
		    \mycaptionfig{Variant effect predictor (VEP) analysis as a function of the number of GWAS catategory number for three different consequences: Missense variant (\textbf{Left}), splice region variant (\textbf{Center}) and 3’ UTR variant (\textbf{Right})}
	    \end{center}
	\end{minipage}

	\vspace{0.9cm} % intersection space

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% FIGURE 2 downstream regulation
	\begin{minipage}{1\columnwidth}
	    \begin{center}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_etissue_y_egene.eps}
		    \end{minipage}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_egene_y_etissue.eps}
		    \end{minipage}
		    \begin{minipage}{0.32\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/pltbar_x_per_variant_egene_etissue_y_gwas.eps}
		    \end{minipage}
		    \mycaptionfig{Count of eGenes per variant-tissue (\textbf{Left}), samples of eQTLs (eTissues) per variant-eGene (\textbf{Center}) and GWAS categories per variant-eGene-eTissue (\textbf{Right}).}
	    \end{center}
	\end{minipage}

	\vspace{0.9cm} % intersection space

@@

	\vspace{0.9cm} % intersection space

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% FIGURE 4 Model
	\begin{minipage}{1\columnwidth}
	    \begin{center}
		    \begin{minipage}{1\columnwidth} 
			    \includegraphics[width=\columnwidth]{img_tmp/graphical_summary.eps}
		    \end{minipage}
		    \mycaptionfig{Model of regulatory variant pleiotropy. I have investigated
	three possible mechanisms of pleiotropy. (\textbf{Left}), Pleiotropic variants have more
	eGenes that result in more functions and more phenotypes. This might arise
	from an enrichment of pleiotropic variants in splicing or 3’ UTR regions.
	(\textbf{Center}), I have also found that eGenes of pleiotropic variants are active
	in more etissues which result in more GWAS phenotypes. This might be
	explained from variants being bound by more transcription factors. (\textbf{Right}),
	Triplets of variant-eGene-eTissues are associated with more GWAS pheno-
	types, which directly affect the number of GWAS phenotypes. I have found
	that this might be explained by an enrichment of missense alleles.}
	    \end{center}
	\end{minipage}

	\vspace{0.9cm} % intersection space

@@

	\begin{itemize}	
	\item We have developed a pipeline of eQTL/GWAS colocalization based on 127 eQTL studies and 150 GWAS (\textbf{Fig. 1}).
	\item These colocalization results are available in an easy to use web site (\textbf{Fig. 1}).
	\item We have computed a list of pleiotropic eQTLs (\textbf{Table 1}).
	\item eGenes of pleiotropic variants are enriched in immune related functions (\textbf{Fig. 2}).
	\item Pleiotropic variants are enriched in missense and splice region VEP effect consequences (\textbf{Fig. 3}).
	\item Pleiotropic eQTLs have more eGenes per variant-eTissue, are active in more eTissues per variant-eGene and are associated to more GWAS categories per variant-eGene-eTissue (\textbf{Fig. 4}).
	\item Pleiotropic variants bind more transcription factors (\textbf{Fig. 5}).
	\item We propose a model explaining the hypotheses underlying the pleiotropy for eQTLs (\textbf{Fig. 6}).
	\end{itemize}
